---
title: "DQ report - MIRACUM"
output:
  html_document:
    df_print: paged
---


```{r include=FALSE}
### LIBRARIES ------------------------
library(knitr)
library(dataquieR)
library(data.table)
library(lubridate)

```

```{r include=FALSE}
### READ DATA -------------------

if(!file.exists("Ergebnisse\\Kohorte.csv") || !file.exists("Ergebnisse\\Observations.csv") ){
  write("Ergebnisse nicht gefunden.", file ="errors/error_message.txt")
  stop("Results not found - aborting.")
}

# read result csv
cohort_report <- read.csv("Ergebnisse\\Kohorte.csv", sep = ";", colClasses = "character")
observations_report <- read.csv("Ergebnisse\\Observations.csv", sep = ";", colClasses = "character")

# read metadata
cohort_metadata <- read.csv("DQ_metadata\\metadata_cohort.csv")
missings <- read.csv("DQ_metadata\\missing-codes.csv")
cohort_contradictions <- read.csv("DQ_metadata\\contradictions_cohort.csv") 
observations_metadata <- read.csv("DQ_metadata\\metadata_observations.csv")

### apply dataquieR format  -------------------

# select columns supported with metadata from cohort
if(!all( cohort_metadata$VAR_NAMES %in% colnames(cohort_report) )){
  write("Metadaten passen nicht zur Kohorte.", file ="errors/error_message.txt")
  stop("Metadata does not match Cohort - aborting.")
}
cohort_report <- cohort_report[,cohort_metadata$VAR_NAMES]

if(!all( observations_metadata$VAR_NAMES %in% colnames(observations_report) )){
  write("Metadaten passen nicht zur Kohorte.", file ="errors/error_message.txt")
  stop("Metadata does not match Cohort - aborting.")
}
observations_report <- observations_report[,observations_metadata$VAR_NAMES]

#extract known data types
cohort_datatyps <- as.data.table(t(cohort_metadata$DATA_TYPE))
colnames(cohort_datatyps) <- cohort_metadata$VAR_NAMES
observations_datatyps <- as.data.table(t(observations_metadata$DATA_TYPE))
colnames(observations_datatyps) <- observations_metadata$VAR_NAMES

# for distubuted analysis value labels or not know 

# function that returns all value labels for distributed dq report
get_labels <- function(vector) {
  
  # unify casing, trim whitespace
  vector <- tolower(vector)
  vector <- trimws(vector, which = "both")
  
  labels <- levels(factor(vector))

  r <- ""
  i <- 1
  for (l in labels) {
    r <- paste0(r,i," = ",l," | ")
    i <- i + 1
  }
  r <- paste0(r,i," = other")
  
  return(r)
}

# find value labels
cohort_metadata$VALUE_LABELS[2] <- get_labels(cohort_report$gender)
cohort_metadata$VALUE_LABELS[7] <- get_labels(cohort_report$icd)
cohort_metadata$VALUE_LABELS[8] <- get_labels(cohort_report$system)
cohort_metadata$VALUE_LABELS[9] <- get_labels(cohort_report$patient_type_fhir_class)
cohort_metadata$VALUE_LABELS[10] <- get_labels(cohort_report$rank)
cohort_metadata$VALUE_LABELS[11] <- get_labels(cohort_report$diagnosis_use)
cohort_metadata$VALUE_LABELS[12] <- get_labels(cohort_report$icd_family)
cohort_metadata$VALUE_LABELS[13] <- get_labels(cohort_report$rank_indicator)
observations_metadata$VALUE_LABELS[2] <- get_labels(observations_report$loinc_code)
observations_metadata$VALUE_LABELS[3] <- get_labels(observations_report$system)
observations_metadata$VALUE_LABELS[4] <- get_labels(observations_report$display)
observations_metadata$VALUE_LABELS[6] <- get_labels(observations_report$unit)

# extract value labels
cohort_value_labels <- as.data.table(t(cohort_metadata$VALUE_LABELS))
colnames(cohort_value_labels) <- cohort_metadata$VAR_NAMES
observations_value_labels <- as.data.table(t(observations_metadata$VALUE_LABELS))
colnames(observations_value_labels) <- observations_metadata$VAR_NAMES
# function that returns only labels in string as list
get_labels <- function(string) {
  # regex to split at | num = : \s?\|?\s?\d\s\=\s
  # [-1] removes 1 = at start
  return(strsplit(string, "\\s?\\|?\\s?\\d\\s\\=\\s")[[1]][-1])
}
cohort_value_labels <- sapply(cohort_value_labels, get_labels)
observations_value_labels <- sapply(observations_value_labels, get_labels)

#set up missing codes
missing_unknown <- missings$CODE_VALUE[2]
#item has integer/string type but cannot be converted to date for age calculation
missing_datetime <- missings$CODE_VALUE[3]
#item has correct type but is missing in value labels
missing_label <- missings$CODE_VALUE[4] 


### Transform to dataquiR format -------------------

# function that sets integer values according to labels and missing codes
apply_labels <- function(vector, labels) {
  
  # unify casing, trim whitespace
  vector <- tolower(vector)
  labels <- tolower(labels)
  vector <- trimws(vector, which = "both")
  labels <- trimws(labels, which = "both")
  
  r <- NULL
  for(item in vector) {
    # default assumes value of item is missing in value labels
    value <- missing_label
    # if item is whitespace or NA set value to NA for correct dataquiR missingness
    if(item == "" || is.na(item) ) value <- NA  
    else {
      i <- 0
      for(l in labels) {
        i <- i + 1
        # if item is found in labels set value to corresponding integer
        if(item == l) value <- i
      }
    }
    r <- c(r,value)
  }
  
  return(r)
}

calculate_age <- function(vector) {

  # ws is assinged missing code for wrong format as well  
  # # unify casing, trim whitespace
  # vector <- tolower(vector)
  # vector <- trimws(vector, which = "both")
  
  #convert to datetime with at least years (YYYY or any YYYY-MM... format)
  r <- parse_date_time(vector, "ymd", truncated = 2)
  
  # calculate age with lubridate period
  r <- lapply(r, function(x) {as.period(interval(x, Sys.Date()), unit = "year")$year} )
  r <- unlist(r)
  
  for (i in 1:length(r) ) {
    if( is.na( r[i] ) && !is.na(vector[i]) ) r[i] <- missing_datetime
  } 
  
  return(r)
}

#apply labels and fix typing per column - missmatched typing is returned as NA (NOT as missing code)

# cohort
cohort_report$birthdate <- as.integer(calculate_age(cohort_report$birthdate))
cohort_report$gender <- as.integer(apply_labels(cohort_report$gender,cohort_value_labels$gender))
cohort_report$patient_zip <- as.integer(cohort_report$patient_zip)
cohort_report$admission_date <- parse_date_time(cohort_report$admission_date, "ymd", truncated = 2)
cohort_report$discharge_date <- parse_date_time(cohort_report$discharge_date, "ymd", truncated = 2)
cohort_report$recorded_date <- parse_date_time(cohort_report$recorded_date, "ymd", truncated = 2)
cohort_report$icd <- as.integer(apply_labels(cohort_report$icd,cohort_value_labels$icd))
cohort_report$system <- as.integer(apply_labels(cohort_report$system,cohort_value_labels$system))
cohort_report$patient_type_fhir_class <- as.integer(apply_labels(cohort_report$patient_type_fhir_class,cohort_value_labels$patient_type_fhir_class))
cohort_report$rank <- as.integer(apply_labels(cohort_report$rank,cohort_value_labels$rank))
cohort_report$diagnosis_use <- as.integer(apply_labels(cohort_report$diagnosis_use,cohort_value_labels$diagnosis_use))
cohort_report$icd_family <- as.integer(apply_labels(cohort_report$icd_family,cohort_value_labels$icd_family))
cohort_report$rank_indicator <- as.integer(apply_labels(cohort_report$rank_indicator,cohort_value_labels$rank_indicator))

# observations
observations_report$effective_date <- parse_date_time(observations_report$effective_date,"ymdHMS",truncated = 5) #includes hours minutes and seconds, but only years would be mapped to YYYY-01-01
observations_report$loinc_code <- as.integer(apply_labels(observations_report$loinc_code,observations_value_labels$loinc_code))
observations_report$system <- as.integer(apply_labels(observations_report$system,observations_value_labels$system))
observations_report$display <- as.integer(apply_labels(observations_report$display,observations_value_labels$display))
observations_report$value <- as.double(observations_report$value)
observations_report$unit <- as.integer(apply_labels(observations_report$unit,observations_value_labels$unit))

# verify correct typing
#sapply(cohort_report, class)

```

```{r include=FALSE}
### GENERATE REPORTS ----------------------
#import missing list

cohort_dq_report <- dq_report(study_data = cohort_report,
                          meta_data  = cohort_metadata,
                          check_table = cohort_contradictions,
                          MISSING_LIST = missings,
                          label_col  = LABEL)

observations_dq_report <- dq_report(study_data = observations_report,
                          meta_data  = observations_metadata,
                          check_table = cohort_contradictions,
                          MISSING_LIST = missings,
                          label_col  = LABEL)

```

## Data Quality report for the 6. Projectathon of MII: MIRACUM

#### This report was created with the [dataquieR](https://dataquality.ship-med.uni-greifswald.de/tutorials2.html) Package for R.

## Result of contradiction checks:  
```{r message=FALSE, warning=FALSE,echo=FALSE}

cohort_dq_report$long_format$con_contradictions$results[[1]]$SummaryPlot

```

### Contradiction checks as Table:

```{r message=FALSE, warning=FALSE,echo=FALSE}

cohort_dq_report$long_format$con_contradictions$results[[1]]$SummaryTable

```

# Information about distribution:  

## Gender:  
#### Value Labels:  
```{r message=FALSE, warning=FALSE,echo=FALSE}
cohort_metadata$VALUE_LABELS[cohort_metadata$LABEL == "gender"]
cohort_dq_report$long_format$con_limit_deviations$results[[1]]$SummaryPlotList$gender

```

## Age:  
#### Age was calculated from birthdates in YYYY... format.
```{r message=FALSE, warning=FALSE,echo=FALSE}

cohort_dq_report$long_format$con_limit_deviations$results[[1]]$SummaryPlotList$age

```

## ICD10 Codes:  
#### Value Labels:
```{r message=FALSE, warning=FALSE,echo=FALSE}
cohort_metadata$VALUE_LABELS[cohort_metadata$LABEL == "icd"]
cohort_dq_report$long_format$con_limit_deviations$results[[1]]$SummaryPlotList$icd

```

## Patient Type FHIR Class:  
#### Value Labels:
```{r message=FALSE, warning=FALSE,echo=FALSE}
cohort_metadata$VALUE_LABELS[cohort_metadata$LABEL == "patient_type_fhir_class"]
cohort_dq_report$long_format$con_limit_deviations$results[[1]]$SummaryPlotLis$patient_type_fhir_class

```

## Observation Value:
```{r message=FALSE, warning=FALSE,echo=FALSE}
observations_dq_report$long_format$con_limit_deviations$results[[1]]$SummaryPlotList$value

```

## LOINC Codes:
Observation Vlaue includes measurements for different LOINC Codes, for a more detailed plausibility/limit check we need to know what LOINC codes are found in the distributed analysis.
```{r message=FALSE, warning=FALSE,echo=FALSE}
observations_metadata$VALUE_LABELS[observations_metadata$LABEL == "loinc_code"]
observations_dq_report$long_format$con_limit_deviations$results[[1]]$SummaryPlotList$loinc_code

```

## Units:
```{r message=FALSE, warning=FALSE,echo=FALSE}
observations_metadata$VALUE_LABELS[observations_metadata$LABEL == "unit"]
observations_dq_report$long_format$con_limit_deviations$results[[1]]$SummaryPlotList$unit

```


# Information about completness: 

### for Cohort:

```{r message=FALSE, warning=FALSE,echo=FALSE,out.width="100%"}

cohort_dq_report$long_format$com_item_missingness$results[[1]]$SummaryTable

```

### for Observations
```{r message=FALSE, warning=FALSE,echo=FALSE,out.width="100%"}

observations_dq_report$long_format$com_item_missingness$results[[1]]$SummaryTable

```

# Plot of completness with missing codes:
#### MISSING 9991100: Missing - reason unknown
#### MISSING 9991200: Deleted - Datetime format not supported
#### MISSING 9991300: Deleted - not in VALUE_LABELS
Regarding completeness and missing codes:   
If a field has the wrong datatype it is evaluated to NA and tracked under Sysmiss in the completeness table. If a field has the correct typing but an undefined value (for example any string as a gender other than 'male', 'female', 'other', 'unknown') the missing code 9991300 is used instead.
```{r message=FALSE, warning=FALSE,echo=FALSE,out.width="100%"}

#Plot
cohort_dq_report$long_format$com_item_missingness$results[[1]]$SummaryPlot
observations_dq_report$long_format$com_item_missingness$results[[1]]$SummaryPlot

```

# Further possible analysis with dataquieR:    
```{r message=FALSE, warning=FALSE,echo=FALSE,out.width="100%"}

cohort_dq_report$app_mat$ApplicabilityPlot
observations_dq_report$app_mat$ApplicabilityPlot

```